---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
To begin the data must be downloaded in to the working directory.
``` {r download}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", 
              destfile = "data.zip")
unzip("data.zip")
```
The data was saved as activity.csv. The following code will then store the data into a data frame.
``` {r read}
actData <- read.csv("activity.csv", header = TRUE, 
                    colClasses = c("integer", "Date", "integer"))
```

## Mean and median steps
Since the the data contains missing values, they will be ignored for now.
```{r remove}
noNa <- na.omit(actData)
```
### Daily steps
A vector of daily steps can be created with the following code.
```{r daily}
daily <- integer()
dailyList <- split(noNa, noNa$date)
## Iterate over each day and put the sum of steps into daily vector
for(i in 1:53) {
    daily[i] <- sum(data.frame(dailyList[i])[, 1])
}
```
Plotting the daily steps yields the following histogram:
``` {r histogram}
hist(daily, xlab = "Daily Steps", main = "", col = "red", breaks = 10)
```

The mean and median steps are calculated and displayed.
``` {r mm,}
mn <- mean(daily)
md <- median(daily)
mm <- c(Mean = mn, Median = md)
print(mm, type = "html")
```

## Average daily activity patterns
A data frame for average step for each interval is made as followed.
``` {r iterval}
dInterval <- aggregate(noNa$steps, by = list(noNa$interval), mean)
```
A time series plot of these is generated:
```{r timeSeries}
plot(dInterval, type = "l", xlab = "Interval", ylab = "Mean Steps")
```  

  The interval with the maximum number of steps is shown.
```{r}
for(i in 1:length(dInterval[, 2])){
    if(dInterval[i, 2] == max(dInterval[, 2])) {
        maxInterval <- dInterval[i,1]}
}
maxInterval
```

## Imputing missing values



## Differences in activity patterns between weekdays and weekends
